Molecular Gas \old{Dynamics} \new{Kinematics and Star Formation Properties} of the Strongly-Lensed Quasar Host Galaxy RXS J1131$-$1231

### 2016-12-13 WRITEUP/diff.tex
change the tone on feedback

\input{preamble}
\slugcomment{To be submitted to the ApJ}
 \makeatletter
% \renewcommand\normalsize{\@setfontsize\normalsize\@xpt{11.56}}
\renewcommand\normalsize{\@setfontsize\normalsize{10.56}{11.4}}
\makeatother

@@

\title{Molecular Gas \old{Dynamics} \new{Kinematics and Spectral Energy Distribution Properties} of the Strongly-Lensed Quasar Host Galaxy RXS J1131$-$1231}
\author{T. K. Daisy Leung, Dominik A. Riechers, and Riccardo Pavesi}
\affil{Department of Astronomy, Space Sciences Building, Cornell University,
Ithaca, NY 14853, USA; \myemail}

@@

\begin{abstract}
We report observations of \bco and \cco line emission towards
the quadruply-lensed quasar
RXS\,J1131$-$1231 at $z$\,=\,0.654 obtained using the
\pdbi (PdBI) and the \carma (CARMA).
Our lens modeling shows that the asymmetry in the double-horned \bco line profile
is mainly a result of differential lensing, where
the magnification factor varies from $\sim$3 to $\sim$9 across
different kinematic components.
The intrinsically symmetric line profile with a conspicuous source-plane velocity
gradient suggests that
the host galaxy is an extended rotating disk, with a CO size of $R_{\rm CO}$$\sim$6\,kpc and
a dynamical mass of $M_{\rm dyn}$\ssim8\E{10}\,\Msun.
In addition, we find a secondary CO-emitting source near RXS\,J1131$-$1231 whose location is
consistent with the optically-faint companion reported in previous studies.
The lensing-corrected molecular gas masses are $M_{\rm gas}$\eq(1.4\pmm0.3)\E{10}\,\Msun
and (2.0\pmm0.1)\E{9}\,\Msun for RXS\,J1131$-$1231 and the companion, respectively.
The gas mass fraction of \old{$\sim$19\% } \new{$\sim$18\pmm4\%} is consistent with the trend of decreasing molecular gas fraction since $z$\ssim2 reported in recent studies.
\old{We find lensing-corrected stellar and dust masses, an IR luminosity, and a
star formation rate of
\mbox{$M_{*}$\eq(3\pmm1)\E{10}\,\Msun}, \mbox{$M_{\rm dust}$\ssim3\E{8}\,\Msun}, \mbox{$L_{\rm IR}$\ssim1.5\E{12}\,\Lsun}, and \mbox{SFR$_{\rm FIR}$\eq(120\pmm63)\,\sfrU} for RXS\,J1131$-$1231.
The source-plane morphologies of its gas and stellar populations
indicate that the star-formation activity may be as extended as the molecular gas reservoir.
We find a \mbox{$M_{\rm BH}$$/$$M_{\rm bulge}$} ratio of \mbox{$>$0.27\petm{0.11}{0.08}\%} which is higher
than the
local value, suggesting that the bulk of the black hole mass of RXSJ1131-1231 is largely in place while its stellar bulge is still assembling.}
\new{We find a lensing-corrected stellar mass of
$M_{*}$\eq(3\pmm1)\E{10}\,\Msun, a star formation rate of SFR$_{\rm FIR}$\eq(120\pmm63)\,\sfrU, and
a star formation efficiency and specific SFR comparable to non-quasar host galaxies at $z$\ssim1.
This together with a gas mass fraction consistent with the trend of evolution indicate 
that no quenching of star formation is seen in RXJ1131.
Our results therefore suggest that 
the decrease in molecular gas mass fraction towards lower redshift 
plays a dominant role 
in the decline of cosmic star formation history even in quasar host galaxies.}
\end{abstract}

@@

The VLA C-band continuum image in \Fig{cont} shows resolved emission from the
jets and the core of the foreground elliptical galaxy
as well as emission toward the background quasar.
Multiple peaks are seen along the arc with their centroids
coincident with the optical emission from the quasar.
We extract the flux densities for the lensing arc and the radio core in \Tab{photometry}.
\old{We find a spectral index of $\alpha^{\rm 2mm}_{\rm 6cm}$\,=\,$-$0.024
for the foreground
galaxy and $\alpha^{\rm 2mm}_{\rm 6cm}$\,=\,$-$0.345
for the background galaxy by fitting a
power-law ($S_\nu \propto \nu^{\alpha}$) to the continuum emission at
5\,GHz and 2\,mm.}
\new{We find a spectral index of $\alpha^{\rm 2mm}_{\rm 6cm}$\,=\,$-$0.02\pmm0.07
for the foreground galaxy and $\alpha^{\rm 2mm}_{\rm 6cm}$\,=\,$-$0.35\pmm0.21 % differential
for the background galaxy by fitting
power-laws ($S_\nu \propto \nu^{\alpha}$) to their continuum fluxes at
5\,GHz and 2\,mm. 
The spectral slope derived for the background source is flatter than the typical slope of pure synchrotron emission \mbox{\citep[$\alpha$\ssim$-$0.7; e.g.,][]{Andreani93a}}. This likely suggests 
that at least a fraction of the observed 2\,mm emission arises from thermal dust emission. 
This spectral slope would be even shallower 
if the background source contributes to the unresolved fraction of the
2\,mm flux. 
In this case, the 2\,mm flux of the foreground galaxy would be lower than the value reported here and yield a steeper slope.
This would also explain the flat spectrum found for the foreground galaxy, which is atypical of elliptical galaxies. 
Assuming a spectral slope of $\alpha$\ssim$-$0.7 for synchrotron radiation in RXJ1131, we expect
a flux density of $S_{\rm 2mm}$\eq0.122\pmm0.004 mJy at 2\,mm. 
The flux excess of $S_{\rm 2mm}$\eq0.27\pmm0.08\,mJy therefore likely arises due to thermal dust emission.
}

@@

The specific star formation rate (sSFR\eq SFR/$M_*$) of 4\petm{2.6}{2.4}\,Gyr\pmOne derived for RXJ1131
is $\lesssim$1.5$\sigma$ above the main sequence according to
the redshift-dependent ``main sequence'' relation in \citet[and references therein]{Tacconi13a}.
Given that RXJ1131 share similar \SF rate, \SF efficiency, and CO disk size as other ``main sequence'' disk galaxies,
the small elevation in sSFR over the main sequence at $z$\ssim0.7 suggests
that the \SF activity in RXJ1131 may be enhanced by interactions with the companion.
\new{In addition, a SFE and a sSFR similar to those of high-$z$ disk 
galaxies, which are not known to host an AGN,
together with a molecular gas mass fraction that follows the reported trend of evolution (\Sec{frac})
suggest that no quenching of star formation is observed in RXJ1131, despite 
that fact that it is hosting a quasar.
Hence, the decline in star formation history since $z$\ssim2 even for AGN host galaxies is likely
a manifestation of the evolution in the molecular gas content rather than negative feedback driven by their central AGNs.
% The decrease in molecular gas mass fraction may be due to 
% decrease in intergalactic cold gas accretion (flows of cold gas/cold gas accretion from the IGM).
}

@@

\old{One explanation for the offset may arise from the fact that BLRs are gravitationally bound to
the central AGN, whereas this condition is not necessary for NLRs.
From an SDSS sample of $>$3800 quasars, \mbox{\citet{Richards02a}} find a
median of \mbox{97\kms} and a dispersion of \mbox{269\,\kms} in the distribution of
the offset between [Mg\,{\scriptsize II}] and [O\,{\scriptsize III}] lines, suggesting
that large offsets are uncommon. Nevertheless,
offsets as large as \mbox{$\sim$\,500\,\kms} have been observed in some cases.}
\new{Velocity offsets between broad line region (BLR) and NLR lines have been reported for other 
AGNs in the literature.
\mbox{\citet{Richards02a}} find a median offset of 
% median of 97\kms and a dispersion of 269\,\kms in the distribution of offset
$\sim$100\pmm270\,\kms 
between [Mg\,{\scriptsize II}] and [O\,{\scriptsize III}] lines in a
sample of $>$3800 quasars,
and \mbox{\citet{Bonning07a}} report
a mean offset of 
% a mean offset of 
% 100\,\kms with a dispersion of 212\,\kms
%FWHM = 500
$\sim$100\pmm210\,\kms
between the broad component of
H$\beta$ and [O\,{\scriptsize III}] lines in a sample of $\sim$2600 quasars at 0.1$<$$z$$<$0.8,
where only $\lesssim$20 of them (i.e., $<$1\%) are found to have offsets $>$800\,\kms and $\sim$1\% 
are found to have offsets $>$500\,\kms}\footnote{\citet{Bonning07a} report the fraction of objects with offset velocities greater 
than 500, 800, 1000, 1500, 2000, and 2500\,\kms. We therefore quote the two fractions corresponding to 
offset velocities closest 
to that of RXJ1131 in this discussion.}.\new{
Thus, large velocity offsets between BLR and NLR lines 
comparable to that of RXJ1131 are uncommon but have been observed in some cases.
}

@@

While properties such as CO linewidth, SFR, and gas mass found in RXJ1131
are consistent with those of local ULIRGs and high-$z$ starburst galaxies,
its SFE is comparable to those of nearby and high-$z$ disk galaxies rather than
\SB systems. This is in good agreement with its disk-like kinematic signatures and its extended molecular gas distribution.
We find a specific \SF rate (sSFR\ssim4 Gyr\pmOne) that is $\lesssim$1.5$\sigma$ higher than those of ``main sequence'' galaxies.
The slight elevation in sSFR over the main sequence suggests that
the on-going star formation activity in RXJ1131 could be enhanced by interactions with the companion.
Recent simulations have illustrated that the disk component of a gas-rich
progenitor galaxy with low SFE can be
retained upon merging since the efficiency at removing angular momentum of the gas via
gravitational torques provided by stellar components is reduced in such a system \citep{Springel05a, Robertson06a, Hopkins09a}.
As such, the extended gas disk of RXJ1131 together with its low SFE may indicate
that the \SF in RXJ1131 could form a
larger stellar bulge in the remnant disk galaxy upon coalescing.
\new{This picture is in agreement with that based on the \mbox{$M_{\rm BH}$$-$$M_{\rm bulge}$} relation,
where we find an elevated \mbox{$M_{\rm BH}$$/$$M_{\rm bulge}$} ratio of $>$0.27\petm{0.11}{0.08}\% for 
RXJ1131 compared to the local value.
This suggests that the stellar bulge of RXJ1131 is still assembling in order to evolve onto the local relation.}
% The elevated ratio compared to nearby AGN host galaxies is also consistent the  other studies of high-$z$ AGNs
% reported in the literature
% \mbox{\citep[e.g.,][]{Walter04a, Peng06a,McLure06a}}.}

\new{Theoretical studies have suggested that the negative feedback from an AGN may remove a 
large fraction of molecular gas from its host galaxy, thereby quenching its star formation \mbox{\citep[e.g., ][]{DiMatteo05a}}.
In this study, we find that the SFE and sSFR derived for RXJ1131 are comparable to those of 
non-AGN host galaxies at $z$\ssim1$-$1.5, and that the 
derived molecular gas mass fraction is consistent with the expected trend of evolution reported in other studies. 
We thus conclude that no negative feedback from the AGN is seen on the molecular gas content 
and on the star formation activity in RXJ1131, albeit hosting a quasar. % at the resolution of our \obs 
% Therefore, we propose that the decline in star formation history since $z$\ssim2 even for AGN host galaxies is 
% a manifestation of the evolution in the molecular gas content 
% rather than
% due to negative feedback driven by their central AGNs.
This suggests that even for AGN host galaxies, 
the decrease in molecular gas mass fraction towards lower redshift
is the main reason for the decline in cosmic star formation history since $z$\ssim2.}

>>>>

\input{preamble}
\slugcomment{With changes from the referee report}
 \makeatletter
% \renewcommand\normalsize{\@setfontsize\normalsize\@xpt{11.56}}
\renewcommand\normalsize{\@setfontsize\normalsize{10.56}{11.4}}
\makeatother

@@

\title{Molecular Gas \old{Dynamics} \new{Kinematics and Star Formation Properties} of the Strongly-Lensed Quasar Host Galaxy RXS J1131$-$1231}
\author{T. K. Daisy Leung, Dominik A. Riechers, and Riccardo Pavesi}
\affil{Department of Astronomy, Space Sciences Building, Cornell University,
Ithaca, NY 14853, USA; \myemail}

@@

\begin{abstract}
We report observations of \bco and \cco line emission towards
the quadruply-lensed quasar
RXS\,J1131$-$1231 at $z$\,=\,0.654 obtained using the
\pdbi (PdBI) and the \carma (CARMA).
Our lens modeling shows that the asymmetry in the double-horned \bco line profile
is mainly a result of differential lensing, where
the magnification factor varies from $\sim$3 to $\sim$9 across
different kinematic components.
The intrinsically symmetric line profile with a conspicuous source-plane velocity
gradient suggests that
the host galaxy is an extended rotating disk, with a CO size of $R_{\rm CO}$$\sim$6\,kpc and
a dynamical mass of $M_{\rm dyn}$\ssim8\E{10}\,\Msun.
In addition, we find a secondary CO-emitting source near RXS\,J1131$-$1231 whose location is
consistent with the optically-faint companion reported in previous studies.
The lensing-corrected molecular gas masses are $M_{\rm gas}$\eq(1.4\pmm0.3)\E{10}\,\Msun
and (2.0\pmm0.1)\E{9}\,\Msun for RXS\,J1131$-$1231 and the companion, respectively.
The gas mass fraction of \old{$\sim$19\% } \new{$\sim$18\pmm4\%} is consistent with the trend of decreasing molecular gas fraction since $z$\ssim2 reported in recent studies.
\old{We find lensing-corrected stellar and dust masses, an IR luminosity, and a
star formation rate of
\mbox{$M_{*}$\eq(3\pmm1)\E{10}\,\Msun}, \mbox{$M_{\rm dust}$\ssim3\E{8}\,\Msun}, \mbox{$L_{\rm IR}$\ssim1.5\E{12}\,\Lsun}, and \mbox{SFR$_{\rm FIR}$\eq(120\pmm63)\,\sfrU} for RXS\,J1131$-$1231.
The source-plane morphologies of its gas and stellar populations
indicate that the star-formation activity may be as extended as the molecular gas reservoir.
We find a \mbox{$M_{\rm BH}$$/$$M_{\rm bulge}$} ratio of \mbox{$>$0.27\petm{0.11}{0.08}\%} which is higher
than the
local value, suggesting that the bulk of the black hole mass of RXSJ1131-1231 is largely in place while its stellar bulge is still assembling.}
\new{We find a lensing-corrected stellar mass of
$M_{*}$\eq(3\pmm1)\E{10}\,\Msun and a star formation rate of SFR$_{\rm FIR}$\eq(120\pmm63)\,\sfrU,
corresponding to a specific SFR and star formation efficiency that are comparable to non-quasar host 
disk galaxies at $z$\ssim1.
This together with a gas mass fraction following the redshift evolution indicates
that no evidence for quenching of star formation is seen in RXS\,J1131$-$1231.
This is in agreement with the elevated \mbox{$M_{\rm BH}$$/$$M_{\rm bulge}$} ratio of \mbox{$>$0.27\petm{0.11}{0.08}\%} compared to the
local value, suggesting that the bulk of the black hole 
mass of RXSJ1131-1231 is largely in place while its stellar bulge is still assembling.}
\end{abstract}

@@

The VLA C-band continuum image in \Fig{cont} shows resolved emission from the
jets and the core of the foreground elliptical galaxy
as well as emission toward the background quasar.
Multiple peaks are seen along the arc with their centroids
coincident with the optical emission from the quasar.
We extract the flux densities for the lensing arc and the radio core in \Tab{photometry}.
\old{We find a spectral index of $\alpha^{\rm 2mm}_{\rm 6cm}$\,=\,$-$0.024
for the foreground
galaxy and $\alpha^{\rm 2mm}_{\rm 6cm}$\,=\,$-$0.345
for the background galaxy by fitting a
power-law ($S_\nu \propto \nu^{\alpha}$) to the continuum emission at
5\,GHz and 2\,mm.}
\new{We find a spectral index of $\alpha^{\rm 2mm}_{\rm 6cm}$\,=\,$-$0.02\pmm0.07
for the foreground galaxy and $\alpha^{\rm 2mm}_{\rm 6cm}$\,=\,$-$0.35\pmm0.21 % differential
for the background galaxy by fitting
power-laws ($S_\nu \propto \nu^{\alpha}$) to their continuum fluxes at
5\,GHz and 2\,mm. 
The spectral slope derived for the background source is flatter than the typical slope of pure synchrotron emission \mbox{\citep[$\alpha$\ssim$-$0.7; e.g.,][]{Andreani93a}}. This likely suggests 
that at least a fraction of the observed 2\,mm emission arises from thermal dust emission. 
This spectral slope would be even shallower 
if the background source contributes to the unresolved fraction of the
2\,mm flux. 
In this case, the 2\,mm flux of the foreground galaxy would be lower than the value reported here and  
lead to a slope steeper than $\alpha^{\rm 2mm}_{\rm 6cm}$\ssim$-$0.02, which is flatter than 
that typical of elliptical galaxies. 
Assuming a spectral slope of $\alpha$\ssim$-$0.7 to account for synchrotron radiation in RXJ1131, we expect
a flux density of $S_{\rm 2mm}$\eq0.122\pmm0.004 mJy at 2\,mm. 
The flux excess of $S_{\rm 2mm}$\eq0.27\pmm0.08\,mJy therefore likely arises due to thermal dust emission.
}

@@

The specific star formation rate (sSFR\eq SFR/$M_*$) of 4\petm{2.6}{2.4}\,Gyr\pmOne derived for RXJ1131
is $\lesssim$1.5$\sigma$ above the main sequence according to
the redshift-dependent ``main sequence'' relation in \citet[and references therein]{Tacconi13a}.
Given that RXJ1131 share similar \SF rate, \SF efficiency, and CO disk size as other ``main sequence'' disk galaxies,
the small elevation in sSFR over the main sequence at $z$\ssim0.7 suggests
that the \SF activity in RXJ1131 may be enhanced by interactions with the companion.
\new{In addition, a SFE and sSFR comparable to those of high-$z$ disk 
galaxies, which are not known to host AGNs,
together with a molecular gas mass fraction that follows the redshift evolution reported in other studies (\Sec{frac})
suggest that no quenching of star formation is observed in RXJ1131, despite 
that fact that it is hosting a quasar.
Hence, the decline in star formation history since $z$\ssim2 
for at least a subset of AGN host galaxies 
may be mainly
a manifestation of the evolution in the molecular gas content rather than a result of 
the negative feedback driven by their central AGNs.
}

@@

\old{One explanation for the offset may arise from the fact that BLRs are gravitationally bound to
the central AGN, whereas this condition is not necessary for NLRs.
From an SDSS sample of $>$3800 quasars, \mbox{\citet{Richards02a}} find a
median of \mbox{97\kms} and a dispersion of \mbox{269\,\kms} in the distribution of
the offset between [Mg\,{\scriptsize II}] and [O\,{\scriptsize III}] lines, suggesting
that large offsets are uncommon. Nevertheless,
offsets as large as \mbox{$\sim$\,500\,\kms} have been observed in some cases.}
\new{Velocity offsets between broad line region (BLR) and NLR lines have been reported for other 
AGNs in the literature.
\mbox{\citet{Richards02a}} find a median offset of 
% median of 97\kms and a dispersion of 269\,\kms in the distribution of offset
$\sim$100\pmm270\,\kms 
between [Mg\,{\scriptsize II}] and [O\,{\scriptsize III}] lines in a
sample of $>$3800 quasars,
and \mbox{\citet{Bonning07a}} report
a mean offset of 
% a mean offset of 
% 100\,\kms with a dispersion of 212\,\kms
%FWHM = 500
$\sim$100\pmm210\,\kms
between the broad component of
H$\beta$ and [O\,{\scriptsize III}] lines in a sample of $\sim$2600 quasars at 0.1$<$$z$$<$0.8,
where only $\lesssim$20 of them (i.e., $<$1\%) are found to have offsets $>$800\,\kms and $\sim$1\% 
are found to have offsets $>$500\,\kms}\footnote{\citet{Bonning07a} report the fraction of objects with offset velocities greater 
than 500, 800, 1000, 1500, 2000, and 2500\,\kms. We therefore quote the two fractions corresponding to 
offset velocities closest 
to that of RXJ1131 ($\sim$715\,\kms) in this discussion.}.\new{
Thus, large velocity offsets between BLR and NLR lines 
comparable to that of RXJ1131 are uncommon but have been observed in some cases.
}

@@

While properties such as CO linewidth, SFR, and gas mass found in RXJ1131
are consistent with those of local ULIRGs and high-$z$ starburst galaxies,
its SFE is comparable to those of nearby and high-$z$ disk galaxies rather than
\SB systems. This is in good agreement with its disk-like kinematic signatures and its extended molecular gas distribution.
We find a specific \SF rate (sSFR\ssim4 Gyr\pmOne) that is $\lesssim$1.5$\sigma$ higher than those of ``main sequence'' galaxies.
The slight elevation in sSFR over the main sequence suggests that
the on-going star formation activity in RXJ1131 could be enhanced by interactions with the companion.
Recent simulations have illustrated that the disk component of a gas-rich
progenitor galaxy with low SFE can be
retained upon merging since the efficiency at removing angular momentum of the gas via
gravitational torques provided by stellar components is reduced in such a system \citep{Springel05a, Robertson06a, Hopkins09a}.
As such, the extended gas disk of RXJ1131 together with its low SFE may indicate
that the \SF in RXJ1131 could form a
larger stellar bulge in the remnant disk galaxy upon coalescing.
\new{This picture is in agreement with the one based on the \mbox{$M_{\rm BH}$$-$$M_{\rm bulge}$} relation,
where we find an elevated \mbox{$M_{\rm BH}$$/$$M_{\rm bulge}$} ratio of $>$0.27\petm{0.11}{0.08}\% for 
RXJ1131 compared to the local value.
This suggests that the stellar bulge of RXJ1131 is still assembling in order to evolve onto the local relation.}
% The elevated ratio compared to nearby AGN host galaxies is also consistent the  other studies of high-$z$ AGNs
% reported in the literature
% \mbox{\citep[e.g.,][]{Walter04a, Peng06a,McLure06a}}.}

\new{Theoretical studies have suggested that negative feedback from an AGN may remove a 
large fraction of the molecular gas from its host galaxy, thereby quenching its star formation \mbox{\citep[e.g., ][]{DiMatteo05a}}.
In this study, we find that the star formation efficiency and specific SFR derived for RXJ1131 are comparable to those of 
non-AGN host disk galaxies at $z$\ssim1$-$1.5, and that the 
derived molecular gas mass fraction is consistent with the redshift evolution reported in other studies. 
We thus find no evidence of negative AGN feedback on the molecular gas content 
and on the star formation activity in RXJ1131 at the resolution of our data.
Our results may be hinting that at least for a subset of AGN host galaxies at intermediate redshift,
their contribution to the decline in cosmic star formation history since $z$\ssim2 
may be mainly
a manifestation of the redshift evolution in the molecular gas content rather than a result of 
negative feedback driven by their central AGNs.
}
